Reproducible Research: Peer Assessment 1
========================================================



### Loading and preprocessing the data
If we do not have the data, we can download it.
```{r echo=TRUE}
# url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
# dest <- "./repdata_data_activity.zip"
# download.file(url, dest)
# unzip(dest)
```

Loading the data (using read.csv()). Missing values are coded as NA. 
```{r echo=TRUE}
data<-read.csv("./activity.csv"
               ,header=TRUE
               ,sep=","
               ,na.strings="NA"
               ,colClasses=c("integer","Date","integer") )
```
The transforming the data into a format suitable for my analysis is not necessary.



### What is mean total number of steps taken per day?

Below you can find a histogram of the total number of steps taken each day
```{r echo=TRUE, fig.width=7, fig.height=6}
help<-as.data.frame(sapply(split(data$steps, data$date), sum))[,1]
hist(help
     ,main="Histogram of the total number of steps taken each day"
     ,xlab="Total number of steps"
     ,col="coral")
```

The code for calculate the mean and median total number of steps taken per day is following:
```{r echo=TRUE}
help_mean<-as.data.frame(sapply(split(data$steps, data$date), mean))[,1]
help_median<-as.data.frame(sapply(split(data$steps, data$date), median))[,1]
```

And the code for report them in graphical form:
```{r echo=TRUE, fig.width=7, fig.height=6}
par(mfrow=c(2,1))
# row 1 col 1
plot(help_mean
     ,x=unique(data$date)
     ,xlab="Date"
     ,ylab=""
     ,main="Mean total number of steps taken per day"
     ,col="green")
# row 2 col 1
plot(help_median
     ,x=unique(data$date)
     ,xlab="Date"
     ,ylab=""
     ,main="Median total number of steps taken per day"
     ,col="blue")
```




### What is the average daily activity pattern?

At the beginnig I splited the steps into interval and uses fuction sapply() to calculate mean for each of interval.
```{r echo=TRUE}
help_interval<-as.data.frame(sapply(split(data$steps, data$interval), mean, na.rm=TRUE))[,1]
```

Afterthat I made a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r echo=TRUE, fig.width=7, fig.height=6}
plot(unique(data$interval)
     ,help_interval
     ,type="l"
     ,main="The average number of steps taken, averaged across all days"
     ,xlab="5-minute interval"
     ,ylab=""
     ,col="brown")
```

The following line of code shows us which 5-minute interval, on average across all the days in the dataset contains the maximum number of steps.
```{r echo=TRUE}
data[which(data$steps==max(data$steps,na.rm=TRUE)),3]
```



### Imputing missing values
We can calculate and report the total number of missing values in the dataset using two simple lines.
```{r echo=TRUE}
summary(data)
length(data[data$steps=="NA",1])
```
As we see the data has 2304 empty values.

We can fill in these lacks using the mean of all datas for which we know number of steps. Additional I took the floor of the mean, because we fill in the number of steps (it should be integer)
```{r echo=TRUE}
floor(mean(data$steps, na.rm=TRUE))
```

I copied the data to data_new and filled in empty values by the floor of mean.
```{r echo=TRUE}
data_new<-data
data_new[is.na(data_new$steps),1]<-floor(mean(data$steps, na.rm=TRUE)) 
```

At the end we can show the histogram of the total number of steps taken each day. For this I created the helpful dataset.
```{r echo=TRUE}
help_new<-as.data.frame(sapply(split(data_new$steps, data_new$date), sum))[,1]
hist(help_new
     ,main="Histogram of the total number of steps taken each day"
     ,xlab="Total number of steps"
     ,col="aquamarine4")
```

We can compare two dataset using summary().
```{r echo=TRUE}
summary(data$steps)
summary(data_new$steps)
```
The mean differs a little, because I filled in them the floor of mean. The 3rd Qu is higher, because I filled empty values the floor of mean.






### Are there differences in activity patterns between weekdays and weekends?
I created a new factor variable with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day and linked this with our data.
```{r echo=TRUE}
part_week<-rep("weekday", length(data_new$date))
for (i in 1:length(data_new$date)) {
  if (weekdays(data_new$date[i]) == "sobota" || weekdays(data_new$date[i]) == "niedziela") {
    part_week[i]<-"weekend"
  }
}
data_new<-cbind(data_new,part_week)
```

To make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis) I created some helpful datasets. I took only rows with "weekend" or "weekday", splited them by interval and calculated mean.
```{r echo=TRUE}
help_weekend<-data_new[data_new$part_week=="weekend",]
help_Weekend_mean<-as.data.frame(sapply(split(help_weekend$steps, help_weekend$interval), mean))[,1]
help_weekday<-data_new[data_new$part_week=="weekday",]
help_Weekday_mean<-as.data.frame(sapply(split(help_weekday$steps, help_weekday$interval), mean))[,1]
```

The following code creates the plot using the basic plotting system.
```{r echo=TRUE, fig.width=7, fig.height=6}
par(mfrow=c(2,1))
# row 1 col 1
plot(help_Weekend_mean
     ,x=unique(help_weekend$interval)
     ,xlab="Interval"
     ,ylab=""
     ,main="Mean total number of steps taken at weekends"
     ,col="mediumvioletred"
     ,type="l")
# row 2 col 1
plot(help_Weekday_mean
     ,x=unique(help_weekday$interval)
     ,xlab="Interval"
     ,ylab=""
     ,main="Mean total number of steps taken at weekdays"
     ,col="khaki"
     ,type="l")
```