Reproducible Research: Peer Assessment 1
========================================================



### Loading and preprocessing the data
If we do not have the data, we can download it.
```{r echo=TRUE}
# url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
# dest <- "./repdata_data_activity.zip"
# download.file(url, dest)
# unzip(dest)
```

Loading the data (using read.csv()). Missing values are coded as NA. 
```{r echo=TRUE}
data<-read.csv("./activity.csv"
               ,header=TRUE
               ,sep=","
               ,na.strings="NA"
               ,colClasses=c("integer","Date","integer") )
```
The transforming the data into a format suitable is following:
```{r echo=TRUE}
data_steps <- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)
data_interval <- aggregate(steps ~ interval, data = data, mean, na.rm = TRUE)
```
I created two sets, the first is sum of steps divided into date and the second - mean of steps divided into interval.



### What is mean total number of steps taken per day?

Below you can find a histogram of the total number of steps taken each day
```{r echo=TRUE, fig.width=7, fig.height=6}
hist(data_steps$steps
     ,main="Histogram of the total number of steps taken each day"
     ,xlab="Total number of steps"
     ,col="coral")
```

The code for calculate the mean and median total number of steps taken per day is following:
```{r echo=TRUE}
mean(data_steps$steps)
median(data_steps$steps)
```




### What is the average daily activity pattern?

I made a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r echo=TRUE, fig.width=7, fig.height=6}
graphics.off()
plot(data_interval$interval
     ,data_interval$steps
     ,type="l"
     ,main="The average number of steps taken, averaged across all days"
     ,xlab="5-minute interval"
     ,ylab=""
     ,col="brown")
```

The following line of code shows us which 5-minute interval, on average across all the days in the dataset contains the maximum number of steps.
```{r echo=TRUE}
data_interval[which.max(data_interval$steps), ]$interval
```



### Imputing missing values
We can calculate and report the total number of missing values in the dataset using two simple lines.
```{r echo=TRUE}
summary(data)
length(data[data$steps=="NA",1])
sum(is.na(data$steps))
```
As we see the data has 2304 empty values.

We can fill in these lacks using the mean of all datas for which we know number of steps. Additional I took the floor of the mean, because we fill in the number of steps (it should be integer)
```{r echo=TRUE}
floor(mean(data$steps, na.rm=TRUE))
```

I copied the data to data_new and filled in empty values by the floor of mean.
```{r echo=TRUE}
data_new<-data
data_new[is.na(data_new$steps),1]<-floor(mean(data$steps, na.rm=TRUE)) 
```

At the end we can show the histogram of the total number of steps taken each day. For this I created the helpful dataset.
```{r echo=TRUE}
data_steps_new <- aggregate(steps ~ date, data = data_new, sum, na.rm = TRUE)
hist(data_steps_new$steps
     ,main="Histogram of the total number of steps taken each day"
     ,xlab="Total number of steps"
     ,col="aquamarine4")
```

We can compare two dataset using summary().
```{r echo=TRUE}
mean(data_steps_new$steps)
median(data_steps_new$steps)
```







### Are there differences in activity patterns between weekdays and weekends?
I created a new factor variable with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day and linked this with our data.
```{r echo=TRUE}
part_week<-rep("weekday", length(data_new$date))
for (i in 1:length(data_new$date)) {
  if (weekdays(data_new$date[i]) == "sobota" || weekdays(data_new$date[i]) == "niedziela") {
    part_week[i]<-"weekend"
  }
}
data_new<-cbind(data_new,part_week)
```

To make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis) I created some helpful datasets. I took only rows with "weekend" or "weekday", splited them by interval and calculated mean.
```{r echo=TRUE}
steps = aggregate(steps ~ interval + part_week, data_new, mean)
steps_weekend = steps[steps$part_week=="weekend",]
steps_weekday = steps[steps$part_week=="weekday",]
```

The following code creates the plot using the basic plotting system.
```{r echo=TRUE, fig.width=7, fig.height=6}
par(mfrow=c(2,1))
# row 1 col 1
plot(steps_weekend$steps
     ,x=steps_weekend$interval
     ,xlab="Interval"
     ,ylab=""
     ,main="Mean total number of steps taken at weekends"
     ,col="mediumvioletred"
     ,type="l")
# row 2 col 1
plot(steps_weekday$steps
     ,x=steps_weekday$interval
     ,xlab="Interval"
     ,ylab=""
     ,main="Mean total number of steps taken at weekdays"
     ,col="khaki"
     ,type="l")
```